---
title: "Lab 10"
author: "Blue Team: Matt Lawson, Nick Zimmerman "
date: "March 22, 2019"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse); library(stringr); library(forcats); library(lubridate); #library(gridExtra)
```

```{r, message=F, warning=F}
questions <- read_csv("data/Questions_trunc.csv")
answers <- read_csv("data/Answers_trunc.csv")

questions <- questions %>% select(Id:Body) %>% na.omit()
answers <- answers %>% select(Id:Body) %>% na.omit()
```



```{r}
head(questions)
head(answers)
```



## Team



### Matt
__Does the `<code>` tag affect the score of a question or answer?__
```{r}
mattq <- questions %>% mutate(ContainsCode = str_detect(Body, pattern="<code>"))
matta <- answers %>% mutate(ContainsCode = str_detect(Body, pattern="<code>"))

questionMeans <- tribble(
  ~Code, ~Mean, 
  "Yes", colMeans( mattq[mattq$ContainsCode, "Score"], na.rm = TRUE),
  "No", colMeans( mattq[!mattq$ContainsCode, "Score"], na.rm = TRUE)
)
answerMeans <- tribble(
  ~Code, ~Mean, 
  "Yes", colMeans( matta[matta$ContainsCode, "Score"], na.rm = TRUE),
  "No", colMeans( matta[!matta$ContainsCode, "Score"], na.rm = TRUE)
)

 q <- ggplot(data=questionMeans) + 
  geom_bar(aes(x=Code, weight=Mean, fill=Code)) + 
  labs(title="Average Score for Questions", x="Contains Code?", y="Score") + 
   guides(fill=FALSE)
a <- ggplot(data=answerMeans) + 
  geom_bar(aes(x=Code, weight=Mean, fill=Code)) + 
  labs(title="Average Score for Answers", x="Contains Code?", y="Score") +  guides(fill=FALSE)

grid.arrange(q, a, ncol=2)

```

Questions and answers that contain code both have higher scores than questions and answers that do not contain code. The difference is much larger in answers. Questions with code in them receive 6 more points on average than questions without code. Answers with code receive 14 more points on average than answers without code. So, if you want to have a question or an answer with a high score, include some code. This is expected because questions about problems with code should include the code in the question. Similarly, answers about code problems should include the fixed code. Formatting code in the proper "code" section of a stack overflow post makes it very easy for future readers to find and read the code.

### Brandon






### Nick

```{r, warning=FALSE, message=F}

Answers <- read_csv("data/Answers_trunc.csv") 
Answers <- select(Answers, -X7)
Questions <- read_csv("data/Questions_trunc.csv")
Questions <- select(Questions, -X7)

Qs <- select(Questions, Id, CreationDate)
colnames(Qs)[2] <- "QuestionDate"
Ans <- inner_join(Answers, Qs, by = c("ParentId" = "Id"))
Ans <- Ans %>% mutate(responseTime =  difftime(CreationDate,QuestionDate, units = "hours") )
Ans <- Ans %>% mutate(order = 0)

```

```{r, message=F, warning=F}
ggplot(filter(Ans, responseTime > 0 & responseTime < 24))+
  geom_point(mapping = aes(responseTime,Score))
```



### Bing

My features are is how is the score affected when the word API is in the question?

```{r}
  APIquestions <- questions %>% filter(grepl("API",Title))
  APIanswers <- answers %>% filter(grepl("API",Body))
  
  #APIquestions
  #APIanswers
  
  t1 <- APIquestions %>% summarise(type = "questions about API", mean = mean(Score))
  t2 <- APIanswers %>% summarise(type = "answers about API", mean = mean(Score))
  t3 <- questions %>% summarise(type = "all questions", mean = mean(Score))
  t4 <- answers %>% summarise(type = "all answers", mean = mean(Score))
  
  t5 <- bind_rows(t1,t2, t3, t4)
  
  t5
  
  ggplot(data = t5) + geom_col(mapping = aes(x = type, y = mean, fill = type))
  

```

As we can see from our table that questions and answers about API are less than questions and answers in general. In my experience answers on forums about APIs have been hard to understand and implement, so it would make sense that the score is less for them. This can also be attributed to the fact that most people are not looking up and asking questions about APIs, with less exposure it would make it hard for the score for questions and answers about APIs to be more than the general population.