---
title: "Lab 11"
author: 'Blue Team: Matt Lawson,'
date: "4/5/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse); library(babynames); library(forcats); library(stringr)
```

# About `babynames` package
  - Baby name data provided by the SSA (Social Security Administration)
  - Includes names with at least 5 uses
  - Includes year, sex, name, number of occurances, and prop
      - Prop = proportion of people of that sex with that name born in that year

```{r, include=F}
head(babynames, 2)
```

## Individual plots of your name’s popularity over time

```{r}
startWithMat <- filter(babynames, str_detect(babynames$name, "^Mat")) %>% group_by(year) %>% 
  summarize(sumprop = sum(prop))

Matt <- babynames %>% 
  filter(str_detect(name, paste(c( "^Matthew$", "^Mathew$", "^Mat$", "^Matt$"),collapse = '|'))) %>%
  group_by(year) %>% 
  summarize(sumprop = sum(prop))

ggplot() + 
  geom_line(data=startWithMat, mapping=aes(x=year, y=sumprop, color="Starts with 'Mat'"), size=1.25) + 
  geom_line(data=Matt, mapping=aes(x=year, y=sumprop, color="Matt/Matthew"), size=1.25) + 
  labs(x="year", y="Proportion", title="Start with 'Mat' vs. name is 'Matt/Matthew' (or a common spelling)" , color="Type")

nameNowBorn <- Matt %>% filter(year==1998 | year==2017)
startNowBorn <- startWithMat %>% filter(year==2017)


nameDif <- as.numeric(nameNowBorn[2,2] - nameNowBorn[1,2])
nowDif <- as.numeric(nameNowBorn[2,2] / startNowBorn[1,2] )

nameDif*100
nowDif*100
```

My name has dropped in percentage of total names by 0.99% since I was born until now. My name makes up 0.3
My name makes up 44.7% of the names that start with the first three letters.





## Ariel and Rachel regexs

```{r}
ariel <- filter(babynames, str_detect(babynames$name, "Ar[iy]+.l+[^a]?$")) %>%
  filter(year==1973 | year==1988 | year==1990) %>%
  group_by(year, name) %>% summarize(count = sum(n), sumprop = sum(prop))

ariel1973 <- ariel %>% filter(year==1973) %>% nrow()
ariel1988 <- ariel %>% filter(year==1988) %>% nrow()
ariel1990 <- ariel %>% filter(year==1990) %>% nrow()
#############################################################################################################
rachel <- filter(babynames, str_detect(babynames$name, "Ra[ey]?ch.*l")) %>%
  filter(year==1973 | year==1988 | year==1990) %>%
  group_by(year, name) %>% summarize(count = sum(n), sumprop = sum(prop))

rachel1973 <- rachel %>% filter(year==1973) %>% nrow()
rachel1988 <- rachel %>% filter(year==1988) %>% nrow()
rachel1990 <- rachel %>% filter(year==1990) %>% nrow()


arielRachel <- tribble(
  ~Name, ~Year, ~`Unique Spellings`,
  "Ariel", 1973, ariel1973,
  "Ariel", 1988, ariel1988,
  "Ariel", 1990, ariel1990,
  "Rachel", 1973, rachel1973,
  "Rachel", 1988, rachel1988,
  "Rachel", 1990, rachel1990
)
```
<br> </br>
__Number of different ways to spell [Ariel, Rachel] for different years__

```{r, echo=F}
arielRachel
```

<br> </br>
<br> </br>

```{r}
ariel <- filter(babynames, str_detect(babynames$name, "Ar[iy]+.l+[^a]?$")) %>%
  filter(sex=="F" & (year==1973 | year==1988 | year==1990 | year==2017)) %>%
  group_by(year) %>% summarize(count = sum(n), sumprop = sum(prop)) %>% mutate(name="Ariel")

rachel <- filter(babynames, str_detect(babynames$name, "Ra[ey]?ch.*l")) %>%
  filter(sex=="F" & (year==1973 | year==1988 | year==1990 | year==2017)) %>%
 group_by(year) %>% summarize(count = sum(n), sumprop = sum(prop)) %>% mutate(name="Rachel")



total <- ariel %>% rbind(rachel) %>% arrange(year)  %>% group_by(year) %>% summarize(totalProp = sum(sumprop))
names(total)[2] <- "Chance"
```
<br> </br>
__Chances a girl would be named either Rachel or Ariel (Including different spellings) for the given year__

```{r, echo=F}
total
```





## The _Little Mermaid_ Effect

__Did The Little Mermaid cause more baby girls to be named Ariel (and various spellings)?__


```{r, echo=F}
moreAriel <- babynames %>% 
  filter(str_detect(babynames$name, "Ar[iy]+.l+[^a]?$") & babynames$sex == "F") %>% 
  filter(between(year, 1989-10, 1989+10)) %>% 
  group_by(year) %>% 
  summarize(name="Ariel", number = sum(n), totalProp = sum(prop))
# moreAriel

ggplot(data=moreAriel) + 
  geom_line(mapping=aes(x=year, y=number), color="orange") +
  geom_vline(aes(xintercept=1989), color="red") + 
  labs(x="Year", y='Number of babies', title='Baby girls named "Ariel" (or another common spelling)', 
       caption = 'The red line shows the year that "The Little Mermaid" was released.')
```

From the year before the movie released to 2 years after the movie released, the number of baby girls named "Ariel" rose from under 2000 to over 7000. But how did the number of baby girls increase in that time?


```{r, echo=F}
allGirls <- babynames %>% 
  filter(sex=="F") %>%
  filter(between(year, 1989-10, 1989+10)) %>% 
  group_by(year) %>% 
  summarize(number = sum(n), totalProp = sum(prop))

ggplot() + 
  geom_line(data=allGirls, mapping=aes(x=year, y=number), color="blue") +
  geom_vline(aes(xintercept=1989), color="red") +
  labs(x="Year", y='Total Number of babies', title='Number of baby girls born', 
       caption = 'The red line shows the year that "The Little Mermaid" was released.')
  
```

This plot shows that the number of baby girls actually increased at a similar rate. Could this mean that the increase in girls named "Ariel" could be simply from the increase in the number of baby girls?


```{r}
startWithVowel <- babynames %>% 
  filter(str_detect(name, "^[AEIOUaeiou]")) %>%
  filter(between(year, 1988, 1990) & sex=="F") %>%
  group_by(year) %>%
  summarize(name="Starts with vowel", n = sum(n), prop = sum(prop))

ariel <-  babynames %>% 
  filter(str_detect(name, "Ar[iy]+.l+[^a]?$")) %>%
  filter(between(year, 1988, 1990) & sex=="F") %>%
  group_by(year) %>%
  summarize(name="Ariel", n = sum(n), prop = sum(prop))
  
startWithVowel <- rbind(ariel, startWithVowel)
startWithVowel

```




__What was the change in proportion in female Ariel baby names?__


## Your names

```{r}
Matt <- babynames %>% filter(year<=2017 & year>=1979) %>% 
  filter(str_detect(name, paste(c( "^Matthew$", "^Mathew$", "^Mat$", "^Matt$"),collapse = '|'))) %>%
  group_by(year) %>%
  summarize( name = "Matthew", totalProp = sum(prop))
mattBirthYear <- subset(Matt, year == 1998)
ggplot() + 
  geom_line(data = Matt, mapping=aes(x=year, y=totalProp), color="orange") +
  geom_point(dat=mattBirthYear, mapping=aes(x=year, y=totalProp), color="red", size=3) + 
  labs(title="Number of babies named 'Matt' from 1979-2017", y="Proportion", caption="My birth year (1998) in red.")

```


## Nick

```{r}
startsWithNic <- babynames %>% 
  filter(str_detect(babynames$name, "^Nic")) %>%
  group_by(year) %>% summarize(sumprop = sum(prop))

Nick <- babynames %>% 
  filter(str_detect(name, paste(c( "^Ni(c|ck)$", "^Ni(c|ch)olas$"),collapse = '|'))) %>%
  group_by(year) %>% summarize(sumprop = sum(prop))


ggplot()+
  geom_line(data = startsWithNic, mapping = aes(year,sumprop,color = "Starts with Nic")) +
  geom_line(data = Nick, mapping = aes(year,sumprop,color = "Nick")) +
  labs(x="year", y="Proportion", title="Start with 'Nic' vs. name is 'Nick/Nicholas'" , color="Type")

```

This plot shows the proportions of babies named Nick (and spelling variations) and all babies with names that start with the letters 'Nic'.  As you can see they follow the same basic trends, but the larger one is so much larger because there are many female babies that were born that also have names that start with 'Nic' like "Nicole","Nicey", etc. 


## Who Did What

__Matt__

  - Created and setup the team markdown file
  - Created the template for the name popularity over time
  - Completed the Ariel and Rachel Regex section
  - Worked on _The Little Mermaid_ Effect section
  - Created the template for the first plot of the names section
  

<br></br>
<br></br>


__Brandon__

<br></br>
<br></br>



__Bing__

Bing put in what part of the lab you did

<br></br>
<br></br>


__Nick__

Nick put in what part of the lab you did

<br></br>
<br></br>