---
title: "Lab 11"
author: 'Blue Team: Matt Lawson,'
date: "4/5/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse); library(babynames); library(forcats); library(stringr)
```

# About `babynames` package
  - Baby name data provided by the SSA (Social Security Administration)
  - Includes names with at least 5 uses
  - Includes year, sex, name, number of occurances, and prop
      - Prop = proportion of people of that sex with that name born in that year



## Individual plots of your name’s popularity over time
```{r}
Matt1 <- babynames %>% filter(year<=2018 & year>=1998) %>% 
  filter(str_detect(name, paste(c( "^Matthew$", "^Mathew$", "^Mat$", "^Matt$"),collapse = '|'))) %>%
  group_by(year) %>%
  summarize( name = "Matthew", total = sum(n))

ggplot(data=Matt1) +
  geom_bar(aes(x=year, weight=total), fill="orange", color="black") +
  labs(title="Number of babies named 'matt' from 1998-2017", y="number")

#######################################################################
Matt2 <- babynames %>% filter(year>=1979 & year<=1998) %>% 
  filter(str_detect(name, paste(c( "^Matthew$", "^Mathew$", "^Mat$", "^Matt$"),collapse = '|'))) %>%
  group_by(year) %>%
  summarize( name = "Matthew", total = sum(n))


ggplot(data=Matt2) +
  geom_bar(aes(x=year, weight=total), fill="orange", color="black") +
  labs(title="number of babies named 'Matt' from 1979-1998", y="number")

```








## Ariel and Rachel regexs







## The _Little Mermaid_ Effect






## Your names

```{r, include=F}
head(babynames)
```

