---
title: "Lab 9"
author: 'Blue Team: Matt Lawson, Nick Zimmerman'
date: "3/14/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(lubridate)

```


```{r, message=F, warning=F}
trips <- read_csv("Nice_ride_trip_history_2017_season.csv", col_types=cols(
  "Start date"=col_datetime(format="%m/%d/%Y %H:%M"), "End date"=col_datetime(format="%m/%d/%Y %H:%M"), "Account type"=col_factor(levels=c("Member", "Casual"))
))

stations <- read_csv("Nice_ride_2017_Station_Locations.csv")
trips <- na.omit(trips)
```

### What Stations make the most revenue and by which member Account type






### Matt

__Do members or casual riders ride for longer times?__

```{r}
colnames(trips)[colnames(trips)=="Total duration (Seconds)"] <- "time"
colnames(trips)[colnames(trips)=="Account type"] <- "account"

mattdf <- trips %>% group_by(account) %>% summarize(time = mean(time), count = n())
mattdf
```

We can see here that casual users spends just over 5 times as long on average. However, members have riden nearly 2 times more than casual members have.


### Bing

What station has the longest ride times?

```{r}
  colnames(trips)[colnames(trips)=="Total duration (Seconds)"] <- "time"
  colnames(trips)[colnames(trips)=="Start station"] <- "sStation"

  eachStation <- trips %>% group_by(sStation) %>% summarize(time = mean(time)) 
  eachStation
  
  ggplot(data = eachStation) + geom_col(mapping = aes(x= sStation, y = time))
  
  topStations <- eachStation %>% filter(time > 75000)
  
  ggplot(data = topStations) + geom_col(mapping = aes(x= sStation, y = time))
  

```
As we can see from our 2 plots that there are 200 stations, but there are 3 out of the stations that are the most popular and have the longest trips. From our second plot we can see that West Broadway & Logan Ave N, Girard Ave N & W Broadway Ave, and Bryant & West Broadway have the longest rides, showing that these are the stations farthest from the center of the city and the commuters are going the farthest distance.





### Brandon
What Stations docks the most bikes
```{r}
colnames(stations)[colnames(stations)=="Name"] <- "name"
colnames(stations)[colnames(stations)=="Total docks"] <- "docks"

brandondf <- stations %>% group_by(name) %>%
summarize(docks = mean(docks), count = n()) %>%
filter(name == "Hennepin County Government Center" | name == "Lake Street & Knox Ave S" | name == "Coffman Union" | name == "Washington Ave SE & Union Street SE" | name == "100 Main Street SE")

arrange(brandondf, desc(docks))
```

This data shows the top five stations with the most docked bikes.


```{r}
ggplot(data = brandondf) + geom_bar(mapping = aes(x = name, y = docks), stat = "identity")
```


### Nick

There are a total of 19,496 unique trips in this data. 

```{r}
Trips <- trips %>% mutate( trip_id = paste(trips$start_station, trips$end_station, sep = " -> ") )
tripId <- unique(Trips$trip_id)
used <- vector("double",19496)


for(i in seq_along(tripId)){
  used[[i]] <- nrow(filter(Trips, trip_id == i))
}
histo <- tibble(Trip = tripId, Used = used)
histo <- arrange(histo,Used)

ten <- head(histo,10)

ggplot(data = ten)+
  geom_count(mapping = aes(x= Trip))
```






### Who did what



### Matt
Matt worked on his individual section. Matt transformed his coloumn names and then summarized them and displayed the membership versus casual

### Brandon
Brandon worked on his individual porition and the transformed formed his coloumns and also summarized the dock numbers in descending order in order to get the stations with the most docks.









