---
title: "Lab 9"
author: 'Blue Team: Matt Lawson, Nick Zimmerman'
date: "3/14/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(lubridate)

```


```{r, message=F, warning=F}
trips <- read_csv("Nice_ride_trip_history_2017_season.csv", col_types=cols(
  "Start date"=col_datetime(format="%m/%d/%Y %H:%M"), "End date"=col_datetime(format="%m/%d/%Y %H:%M"), "Account type"=col_factor(levels=c("Member", "Casual"))
))

stations <- read_csv("Nice_ride_2017_Station_Locations.csv")
trips <- na.omit(trips)
```

### What docks make the most money and by who

```{r}
colnames(trips)[colnames(trips)=="Start station"] <- "station"
colnames(trips)[colnames(trips)=="Total duration (Seconds)"] <- "time"
colnames(trips)[colnames(trips)=="Account type"] <- "account"
teamdf <- trips %>% group_by(station) %>%  summarize(time = mean(time), count = n()) %>%
filter(time > 75000)
 
```


```{r}
ggplot(data = teamdf) + geom_bar(mapping = aes(x = station, y = time,), stat = "identity")
```

We also wanted to plot the membership on this graph but was not able to do it. But using Matt's data we can conclude the three stations with the most run time and that Casuals spend 5 times more using the bikes therefore, casuals should be most of the income the company makes. Although we are not explicitly told how the money is made, if is based off were you docked is where you payed then looking at Brandon's data you would see the top 5 stations with the most dockage.






### Matt

__Do members or casual riders ride for longer times?__

```{r}
colnames(trips)[colnames(trips)=="Total duration (Seconds)"] <- "time"
colnames(trips)[colnames(trips)=="Account type"] <- "account"

mattdf <- trips %>% group_by(account) %>% summarize(time = mean(time), count = n())
mattdf
```

We can see here that casual users spends just over 5 times as long on average. However, members have riden nearly 2 times more than casual members have.


### Bing

What station has the longest ride times?

```{r}
  colnames(trips)[colnames(trips)=="Total duration (Seconds)"] <- "time"
  colnames(trips)[colnames(trips)=="station"] <- "sStation"

  eachStation <- trips %>% group_by(sStation) %>% summarize(time = mean(time)) 
  eachStation
  
  ggplot(data = eachStation) + geom_col(mapping = aes(x= sStation, y = time))
  
  topStations <- eachStation %>% filter(time > 75000)
  
  ggplot(data = topStations) + geom_col(mapping = aes(x= sStation, y = time))
  

```
As we can see from our 2 plots that there are 200 stations, but there are 3 out of the stations that are the most popular and have the longest trips. From our second plot we can see that West Broadway & Logan Ave N, Girard Ave N & W Broadway Ave, and Bryant & West Broadway have the longest rides, showing that these are the stations farthest from the center of the city and the commuters are going the farthest distance.





### Brandon
What Stations docks the most bikes
```{r}
colnames(stations)[colnames(stations)=="Name"] <- "name"
colnames(stations)[colnames(stations)=="Total docks"] <- "docks"

brandondf <- stations %>% group_by(name) %>%
summarize(docks = mean(docks), count = n()) %>%
filter(name == "Hennepin County Government Center" | name == "Lake Street & Knox Ave S" | name == "Coffman Union" | name == "Washington Ave SE & Union Street SE" | name == "100 Main Street SE")

arrange(brandondf, desc(docks))
```

This data shows the top five stations with the most docked bikes.


```{r}
ggplot(data = brandondf) + geom_bar(mapping = aes(x = name, y = docks), stat = "identity")
```


### Nick

There are a total of 19,496 unique trips in this data. 







### Who did what



### Matt
Matt worked on his individual section. Matt transformed his coloumn names and then summarized them and displayed the membership versus casual

### Brandon
Brandon worked on his individual porition and the transformed formed his coloumns and also summarized the dock numbers in descending order in order to get the stations with the most docks.

### Bing
Bing worked on his individual porition and the team section. Transformed formed his coloumns and also summarized the ride time with the docks with the most ride time.

### Nick
Nick worke don his indivdual portion. Transformed his data by using mutate and a for loop. Nick could not get his code to work








