---
title: "Lab 4"
author: "Blue Team: Matt Lawson, Nick Zimmrman, BLAH, BLAH"
date: "2/5/2019"
output: html_document
---
  
  
    
    
    
    
     
     


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=F}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
babies <- na.omit(babies)
```
```{r, include=F}
head(babies)
```



## Questions and Importance

##### 1. Do mothers who smoke have increased rates of premature delivery (before 270 days)?
  This is an important question because premature delivery can lead to many health problems for the baby. Lung, heart, and brain problems, cerebral palsy, learning imparment, and chronic health issues are some known problems.
  ([www.mayoclinic.org](https://www.mayoclinic.org/diseases-conditions/premature-birth/symptoms-causes/syc-20376730))

##### 2. Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?




## Team Plots
```{r}
ggplot(data=babies, mapping=aes(x=as.factor(Premature)))+
  # geom_bar()+
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  facet_wrap(~smoke)
```



### Recomendation


### Individual Findings


###### Matt

```{r}
ggplot(data=babies) + 
  geom_bar(mapping=aes(x=as.factor(smoke), fill=as.factor(med)), position="dodge") + 
  scale_fill_discrete(name="Mother's Highest Education",
                      breaks=c(0, 1, 2, 3, 4, 5, 7, 9),
                      labels=c("Less than 8th grade", "Some high school", "High school graduate", "High school + trade school", "some college", "College graduate", "trade school", "Unknown")) +
  labs(x="No              Smoker?                Yes")
```


Mothers that have some college experience or a college degree are much less likeley to have ever smoked.
I used geom_bar to plot the mothers education for smokers and for non-smokers. I used color to see the different education leves. I also used position="dodge" to better compare the counts of each diffferent education level. I also changed the legend from numbers to the actual meaning so that the reader doesn't have to reference the help page.

```{r, include=F}
with(babies, table(smoke, Premature))
# 41% of non-smokers are born prematurely, but 52% of smokers are born prematurely 
```




###### Brandan


###### Bing


###### Nick




### Who did what