---
title: "Lab 4"
author: "Blue Team: Matt Lawson, Nick Zimmrman, BLAH, BLAH"
date: "2/5/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=F}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
babies <- na.omit(babies)
```
```{r, include=F}
head(babies)
```



# Questions and Importance

Do mothers who smoke have increased rates of premature delivery (before 270 days)?

Do the newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born)?




# Team Plot


## Recomendation


# Individual Findings


### Matt

```{r}
ggplot(data=babies) + 
  geom_bar(mapping=aes(x=as.factor(smoke), fill=as.factor(med)), position="dodge")
```

Mothers that have never smoked are more likely to have some college experience or a college degree.

```{r}
with(babies, table(smoke, Premature))
# 41% of non-smokers are born prematurely, but 52% of smokers are born prematurely 
```


### Brandan


### Bing


### Nick




# Who did what