---
title: "Lab 13"
author: "Blue Team: Matt Lawson, Bing Mitchell, Brandon Bowen, Nick Zimmerman"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Team Section

__Has average energy consumption per person changed from the 1980s to the 2000s?__

<br></br>

__Description of the Datasets:__

Description goes here















## Bing
Sub Question: Has nuclear energy use increased from 1989 to 2009?

Null Hypothesis: Nuclear Energy use per person hasnt changed in European countries from 1989 to 2009

Alternate Hypothesis: Nuclear Energy use per person has increased in European coutnries from 1989 to 2009

Test Statistic: The average of 2009 - the average of 1989

```{r, echo = FALSE}
  nuclear <- read_csv("nuclear_power_generation_per_person.csv") %>%
gather(2:53, key="year", value="TOEperPerson")  %>% na.omit()

 # TOE per person = amount of electricity produced by nuclear power per person, in tons of oil equivalent
# nuclear
```

permTest Function

it is a perumation test that takes in values for 2 years and gives the mean difference in nuclear use per year 



```{r, echo=TRUE}
permTestBING <- function(x, y, n) {
  output <- vector("double", length(n))
  master <- rbind(x, y)
  nrow(master);nrow(x)
  for (i in c(1:n)) {
    sample <- sample(nrow(master), min(nrow(x), nrow(y)))
    randomX <- master[sample,]
    randomY <- master[-sample,]
    meanX <- mean(randomX$TOEperPerson)
    meanY <- mean(randomY$TOEperPerson)
    
    # print(meanX); print(meanY)
    
    output[[i]] <- meanX - meanY
  }
  return(output)
}
```



```{r, echo = FALSE}
set.seed(999)
nuclear89 <- nuclear %>% filter(year=="1989") %>% select(year, TOEperPerson)
nuclear09 <- nuclear %>% filter(year=="2009") %>% select(year, TOEperPerson)

# nrow(data89); nrow(data09)

testStat <- permTestBING(nuclear89, nuclear09, 1000)
# hist(testStat)

realX <- mean(nuclear89$TOEperPerson)
realY <- mean(nuclear09$TOEperPerson)
percentile <- sum(testStat < (realX-realY))/length(testStat)*100


ggplot() + 
  geom_histogram(mapping=aes(x=testStat), fill="white", color="black") + geom_vline(xintercept=(realX-realY), color="red")


```

We get that the change in Nuclear use from 1989 to 2009 is only the 46th percentile, so we fail to reject our null hypthesis that there hasn't been a change in the nuclear use in the time from 1989 to 2009.



## Matt

$\bf{H_0:}$ There is no difference in average coal consumption in the 1980s and the average coal consumption in the 2000s

$\bf{H_1:}$ There is a significant difference in average coal consumption in the 1980s and the average coal consumption in the 2000s

__Test Statistic:__ Difference in sample means $\bar{X}_x - \bar{X}_y$

<br></br>

__Subquestion: Is there a difference in coal consumption from the 1980s to the 2000s?__

If there is a difference in energy consumption per person from these decades, then maybe there is a difference in coal consumption during the same time.

<br></br>

__Map Function__

```
meanX <- map_dbl(randomX, func);
meanY <- map_dbl(randomY, func)
 ```
 
 This `map_dbl` function takes in a dataframe and applies the function `func` to the columns of it.
 
 <br></br>
 
 __Function with Function Argument__
 
 the function `permTestMATT` below takes in an argument called `func`. `func` gets uses in the map above.


<br></br>
<br></br>
<br></br>

```{r message=FALSE, warning=FALSE}
  coal <- read_csv("coal_consumption_per_cap.csv")  %>%
  gather(2:53, key="year", value="TOEperPerson", convert = TRUE)  %>% na.omit()

 # TOE per person = Coal consuumption per person (in Tons of Oil Equivalency)
coal80 <- coal %>% filter(between(year, 1980, 1989)) %>% select(year, TOEperPerson)
coal00 <- coal %>% filter(between(year, 2000, 2009)) %>% select(year, TOEperPerson)

```

```{r}
permTestMATT <- function(x, y, n, func) {
  set.seed(1)
  output <- vector("double", length(n))
  master <- rbind(x, y)
  for (i in c(1:n)) {
    sample <- sample(nrow(master), min(nrow(x), nrow(y)))
    randomX <- master[sample,]
    randomY <- master[-sample,]

    meanX <- map_dbl(randomX, func)
    meanY <- map_dbl(randomY, func)
    
    output[[i]] <- meanX[[2]] - meanY[[2]]
  }
  return(output)
}
```

<br></br>
<br></br>


```{r warning=FALSE}
statMatt <- permTestMATT(coal80, coal00, 1000, mean)
realX <- mean(coal80$TOEperPerson)
realY <- mean(coal00$TOEperPerson)
percentile <- sum(statMatt < (realX-realY))/length(statMatt)*100
```

<br></br>

The real data is the `r percentile` percentile of the random distribution.

```{r echo=F, message=F, warning=TRUE}
ggplot() + 
  geom_histogram(aes(x=statMatt), color="black", fill="white") + 
  geom_vline(xintercept=(realX-realY), color="red")

```

We can conclude that there is a difference between the average coal consumption per person in the 1980s and the average coal consumption per person in the 2000s. We reject $H_0$ in favor of $H_1$.


## Brandon 

Sub Question: is ther a change of hydro energy consumption per person from 1989 to 2009

Null Hypothesis: There is no signifigant change in energy consumption

Alternative: There is signifigant change in energy consumption

```{r message=FALSE, warning=FALSE}
  hydro <- read_csv("hydro_power_generation_per_person.csv")  %>%
  gather(2:53, key="year", value="HydroperPerson", convert = TRUE)  %>% na.omit()

 # Hydro per person = Hydro consumption per person 
hydro80 <- hydro %>% filter(between(year, 1980, 1989)) %>% select(year, HydroperPerson)
hydro09 <- hydro %>% filter(between(year, 2000, 2009)) %>% select(year, HydroperPerson)

```



```{r}
permTestBrandon <- function(x, y, n, func) {
  set.seed(1)
  output <- vector("double", length(n))
  master <- rbind(x, y)
  for (i in c(1:n)) {
    sample <- sample(nrow(master), min(nrow(x), nrow(y)))
    randomX <- master[sample,]
    randomY <- master[-sample,]

    meanX <- map_dbl(randomX, func)
    meanY <- map_dbl(randomY, func)
    
    output[[i]] <- meanX[[2]] - meanY[[2]]
  }
  return(output)
}
```


```{r warning=FALSE}
stat <- permTestBrandon(hydro80, hydro09, 1000, mean)
realsX <- mean(hydro80$HydroperPerson)
realsY <- mean(hydro09$HydroperPerson)
percentile <- sum(stat < (realsX-realsY))/length(stat)*100
```


```{r echo=F, message=F, warning=TRUE}
ggplot() + 
  geom_histogram(aes(x=stat), color="black", fill="white") + 
  geom_vline(xintercept=(realsX-realsY), color="red")

```




### Nick

Sub question: Has Europe's consumption of oil experienced a significant decrease from 1985 to 2010

Null hypothesis: Consumption hasn't changed

Altenate hypothesis: Consumption has changed
 
Test statistic: Average oil consumption per person in Europe

```{r message=FALSE, warning=FALSE}
library(gapminder)
europe_countries <- gapminder %>% filter(continent == "Europe")
europe_countries <- unique(europe_countries$country)

oil <- read_csv("oil_consumption_per_cap.csv")
oil <- filter(oil, country %in% europe_countries)
oil <- oil %>% gather(2:53, key = "year", value = "Oil/Person", na.rm = TRUE, convert = TRUE)

oil_1985 <- oil %>% filter(between(year, 1985, 1990)) %>% select(year, `Oil/Person`)
oil_2010 <- oil %>% filter(between(year, 2005, 2010)) %>% select(year, `Oil/Person`)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
permTestNick <- function(x, y, n, func){
  set.seed(1)
  output <- vector("double", length(n))
  master <- rbind(x, y)
  for (i in c(1:n)) {
    sample <- sample(nrow(master), min(nrow(x), nrow(y)))
    randomX <- master[sample,]
    randomY <- master[-sample,]

    meanX <- map_dbl(randomX, func)
    meanY <- map_dbl(randomY, func)
    
    output[[i]] <- meanX[[2]] - meanY[[2]]
  }
  return(output)
}

```

```{r warning=FALSE, include=FALSE}
statNick <- permTestNick(coal80, coal00, 1000, mean)
realX_Nick <- mean(oil_1985$`Oil/Person`)
realY_Nick <- mean(oil_2010$`Oil/Person`)
percentile <- sum(statNick < (realX_Nick - realY_Nick))/length(statNick)*100
```

```{r echo=F, message=F, warning=TRUE}
ggplot() + 
  geom_histogram(aes(x=statNick), color="black", fill="white") + 
  geom_vline(xintercept=(realX_Nick - realY_Nick), color="red")

```

