---
title: "Lab3_BlueTeam"
author: "Blue Team"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=F}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

```{r, include=F}
cdc$exerany <- factor(cdc$exerany, labels=c("Not an active excerciser", "Active excerciser"))
cdc$smoke100 <- factor(cdc$smoke100, labels=c("Not a smoker", "Smoker"))
cdc$hlthplan <- factor(cdc$hlthplan, labels=c("No health plan", "Health plan"))
```


```{r, include=F}
head(cdc)
```
### Question

Does smoking affect the weight loss as one gets older? What’s the difference of weight loss between male and female?

### Signifigance
This question is signifigant because if smoking has an adverse effect on weight loss then smokers would genraly become even more unhealthy as they age and have a smaller life span than non-smokers. If smokers have a smaller life span then insurance companies may not want to give smokers the same health plans and benifits as non- smokers. Furthermore, does gender have any effects on weight loss also.

### Results

 As seen from the graph below Smokers and non-smokers experience the same amount or minimal difference of weight loss as they get older in every health class. Also, they have about the same life expectancy. We conclude that smoking has no signifigant effect on weight loss or weight gain and thus no adverse effect on overall health. While females do have an overal less weight than males, this could be due to the difference of male and femal physique.

### Recomendation

Health insurance companies should give the same benifits and health plans to smokers as they do to non-smokers as they don't have any risk factor.




```{r, warning=FALSE, message=F}
ggplot(data = cdc, mapping = aes( x = age, y = weight, color = genhlth)) +
geom_smooth(se = FALSE) +
facet_grid(smoke100~gender) +
  ggtitle('Weight vs. Age broken down by smoking habits and gender')
```

***

### Individual Section

```{r, warning=F, message=F}
ggplot(data = cdc, mapping = aes( x = age, y = weight, color = smoke100)) +
geom_smooth(se = FALSE) +
facet_grid(genhlth~gender)
```

We can see here that age vs. weight is different for male and female. It is also different for different levels of general health. However, it is nearly exactly the same for smokers vs. non smokers.

***


```{r, warning=F, message=F}
ggplot(data = cdc, mapping = aes( x = age, y = weight, color = genhlth)) +
geom_smooth(se=FALSE) +
  facet_wrap(~gender) + 
   scale_color_discrete(name = 'Health Level')
```

We can see here that different health levels age differently with respect to weight. We can also see how men begin to decrease in weight sooner than women do.

***

```{r}
ggplot(data = cdc) +
geom_bar(mapping = aes(x = hlthplan, fill = genhlth), position = "fill") +
ggtitle(" Is the proportion of healthy people without a health plan less than 
the proportion of healthy people with a health plan")
```
   I Brandon Bowen created a bar graph to show people with and without a health plan and the proportion of each health class in each section. To easily see the proportion I made the function "position = "fill". " I choose this over count becuase it made it an easier comparison for my question. Then i used the "fill" function to classify my general health to make it easier to see. Based off the chart in the upper two catagories of health( Exccelent and very good) the proportion with a health plan is larger than the one without. In the lower three catagories the proportion is larger with those who don't have a health plan. Therefore i conclude the most healthy people will genrally have a helath plan instead of those who don't. This can be attributed to the fact that people who are healthier most likely have some form of a health plan.

***

```{r}
ggplot(data=cdc) +
  stat_count(mapping=aes(x=exerany, fill=exerany)) +
  facet_wrap(~genhlth) +
  coord_flip() +
  guides(fill=FALSE) +
  ggtitle("Does excercising have an effect on general health?")
```

Clearly, excercising had a positive effect on general health. The vast majority of the people in "excelent", "very good", and "good" general health are active excercisers.

***

```{r}
ggplot(data=cdc) +
  geom_point(aes(x=height, y=weight))
```

```{r}
ggplot(data=cdc) +
  geom_jitter(aes(x=weight, y=wtdesire))
```

```{r}
ggplot(data = cdc) + 
  geom_jitter(data = cdc, mapping = aes(x = weight, y = age, color = hlthplan)) +
  facet_wrap(~ hlthplan) +
  ggtitle("Age and Weight distribution based on whether or not someone has a healthplan")
```
Age and weight tend to increase if someone has a healtplan. Since smoking has no affect on someone's age or weight loss it should stand to reason that giving them access to a healthplan would help them live longer. 

***

```{r}
ggplot(data=cdc) + 
geom_bar(mapping=aes(x = exerany, fill = genhlth), position = "fill") + labs(title = "Bar Chart of General Health of people who do and don't excersize")
```



***
### Contribution

Brandon created the Bar graph(" Is the proportion of healthy people without a health plan smaller than the proportion of healthy people who do have a health plan." )


